PyTracer: Automatically Profiling Numerical Instabilities in Python

### 2022-10-17 IEEEtran/paper.tex
Remove \remove of first review

\remove{Listings~\mbox{\ref{fig:wrapper_creation}} and~\mbox{\ref{fig:generic_wrapper}} show
    \mbox{\pytracer}'s function instrumentation mechanism.}
For each function to be instrumented, \pytracer's \texttt{loader} class creates a
wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
that can be called directly with the () operator, because they are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning. 
This limitation is expected to have a minor impact in practice since non-writable
classes are usually called from pure Python classes that can be
instrumented. Therefore, \pytracer reports information about the stability of
non-writable classes through the instrumentation of the pure Python
classes that use them.

@@

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} \remove{and \textit{Slowdown MCA}} shows the slowdown within the Fuzzy environment.
Without Fuzzy, \pytracer introduces an average slowdown of $17\times$ for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, \remove{the Fuzzy instrumentation} using MCA
operations (through Fuzzy) instead of the native ones increased the average
slowdown to $37\times$\remove{for RR and $43\times$ for Full MCA}.

@@

Figure~\ref{fig:performance_parsing} shows the postprocessing time and trace
sizes produced by \pytracer. The linear correlation between the log of the size
and the log of the time shows a significant correlation (p-value $\leq 10^{-8}$
\remove{for RR and p-value $\leq 10^{-3}$ for MCA}).
% \tristan{Mention linear correlation with p-value here.}

@@

\remove{
    As we showed in Section~\ref{sec:impact_mca_modes}, using Full MCA mode leads to
    runtime errors that have a considerable impact on the proper functioning of the
    execution, which is unfortunate since these errors do not reflect actual
    numerical errors but are related to perturbations that should not occur.
    Conversely, RR mode is far more conservative since it preserves exact operations
    and is easier to use even though it does not simulate all perturbations.
    Therefore, from our experiments, we recommend the use of RR over Full MCA. More
    research is required to address the issues encountered with Full MCA. 
}

>>>>

For each function to be instrumented, \pytracer's \texttt{loader} class creates a
wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
that can be called directly with the () operator, because they are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning. 
This limitation is expected to have a minor impact in practice since non-writable
classes are usually called from pure Python classes that can be
instrumented. Therefore, \pytracer reports information about the stability of
non-writable classes through the instrumentation of the pure Python
classes that use them.

@@

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} shows the slowdown within the Fuzzy environment.
Without Fuzzy, \pytracer introduces an average slowdown of $17\times$ for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, using MCA
operations (through Fuzzy) instead of the native ones increased the average
slowdown to $37\times$.

@@

Figure~\ref{fig:performance_parsing} shows the postprocessing time and trace
sizes produced by \pytracer. The linear correlation between the log of the size
and the log of the time shows a significant correlation (p-value $\leq 10^{-8}$
).
% \tristan{Mention linear correlation with p-value here.}

@@
